\bf Service Chain Placement in SDNs \thanks{Research supported in part by Network Programming (Neptune) Consortium, Israel.}

### 2017-11-27 src/chain.tex

\maketitle

@@

We show that even the feasibility problem is NP-hard in general
graphs.  Hence we focus on DAGs.  We show that the problem is still
NP-hard in DAGs even for a very simple network, and even if the VNF
specification consists of only one option (i.e., the virtual DAG is a
path).
%
On the other hand, we present an FPTAS for the case where the network
is a DAG.  In addition, based on our FPTAS, we provide algorithms for
instances in which the service chain passes through a bounded number
of vertices whose degree is larger than two.
\end{abstract}

@@

We show that even feasibility is NP-hard in general networks using a
reduction from \textsc{Hamiltonian Path}.
%Hence we focus on DAGs.
We show that the problem is still NP-hard if the network is a DAG even
for a very simple network, and even if the specification consists of a
single option.  Both reductions are from \textsc{Partition}.

On the positive side, we present an FPTAS for the case where the
network is a DAG which is based on a non-trivial dynamic programming
algorithm.
%
Based on our FPTAS, we provide a randomized algorithm for general
networks in which there is an optimal placement whose physical path
consists of at most $k$ vertices whose degree is larger than $2$.  For
example, this can be assumed if all simple paths from $s$ to $t$
contain at most $k$ such vertices.  The algorithm computes a
$(1+\eps)$-approximate placement with high probability using $k! \log
n$ invocations of the FPTAS.
%
We also present a (deterministic) parameterized algorithm that
computes a $(1+\eps)$-approximate placement in time $O(k! \cdot
\text{poly}(n))$, where $k$ is the number of vertices in the network
whose degree is larger than $2$.

@@

\item[Virtual specification:]
%
  The description of a request for a service chain consists of a
  physical source $s \in V$, a physical destination $t \in V$, and a
  directed acyclic graph (DAG) $\calG = (\calV,\calE)$.
%
  The DAG $\calG$ has a source node $\sigma \in \calV$ and a
  destination $\tau \in \calV$. Each node $\alpha \in \calV$ represent
  a VNF that has a processing demand $p(\alpha)$.
%
  Without loss of generality, we assume that $p(\sigma) = p(\tau) =
  0$.  Each edge $\eps \in \calE$ has a bandwidth demand $b(\eps)$.

@@

In this section we consider \scp when the physical network is an
undirected graph.  Recall that even the feasibility version of \scp is
NP-hard in general networks, and therefore we focus on a special case.
%
A vertex $v \in V$ is called \emph{neighborly} if it has more than two
neighbors.
%
First, we assume that there exists an optimal placement whose physical
path consists of at most $k$ neighborly vertices.  For example, this
can be assumed if all simple paths from $s$ to $t$ contain at most $k$
neighborly vertices.  In this case we present a randomized algorithm
that computes an $(1+\eps)$-approximate placement with high
probability whose running time is $O(k! \cdot \text{poly}(n))$.
%
Our second algorithm works under the stronger assumption that there
are $k$ neighborly vertices in the network.  In this case, we present
a deterministic algorithm whose running time is $O(k! \cdot
\text{poly}(n))$.

Our randomized algorithm consists of $k!$ iterations of the
following two phases: an orientation phase that applies a random
orientation to the physical network, and an execution of the FPTAS for
DAGs given in Theorem~\ref{thm:fptas}.  The algorithm finds a
$(1-\eps)$-approximate placement with probability $(1-1/e)$, and the
running time of the algorithm is $O((n+t(n))k!)$, where $t(n)$ is the
time it takes to compute a placement when the physical network is a
DAG.
%
As usual, one may amplify the probability of success using repetition.

@@

\begin{figure}[t]
\centering
\scalebox{0.9}{
  \input{fig-orientation}
}
\caption{Orientation phase: on the left is the physical network, where
  only the heavy nodes are drawn.  The numbers represent the
  permutation and the dashed path represents a path of an optimal
  placement.  The orientated physical network is on the right.  It is
  enough that the internal ordering of the nodes on the dashed path is
  "correct" to ensure that the survival of the optimal placement. }
\label{fig:orientation}
\end{figure}

@@

\begin{theorem}
There exists an algorithm, that given an \scp instance with an optimal
solution that contains at most $k$ neighborly vertices, finds a
$(1+\eps)$-approximate placement with high probability, whose
running time is $O(k! \cdot \text{\rm poly}(n))$.
\end{theorem}
\begin{proof}
Consider an optimal placement with at most $k$ neighborly vertices,
and let $\pi'$ be the permutation it induces on the neighborly
vertices in the physical path.
%
If the random permutation $\pi$ orders the neighborly vertices
correctly, i.e., if $\pi(v) = \pi'(v)$, for every $v \in N$, then the
optimal placement is feasible with respect to $G_\pi$.  Hence, the
FPTAS will find a $(1+\eps)$-approximate placement in $G_\pi$.  This
placement is feasible, and therefore $(1+\eps)$-approximate placement,
with respect to $G$.
%
The permutation $\pi$ agrees with $\pi'$ with probability $1/k!$, and
the probability that $\pi$ agrees with $\pi'$ in $k!$ iterations is $1
- (1 - 1/k!)^{k!} \geq 1 - e^{-1}$.  One may amplify the success
probability to $1 - \inv{n}$ using $\log n$ repetitions.
%
\qed
\end{proof}

In the case where $\abs{N} = k$ we can simply examine the $k!$
permutations of the neighborly vertices.  For each possible
permutation the algorithm applies the graph orientation procedure as
described above, and then executes the FPTAS for DAGs given in
Section~\ref{sub:fptas}.  
Thus the running time of the algorithm is $O(k! \cdot \text{\rm poly}(n))$.
%
We note that permutation enumeration for $k$ numbers takes $O(k!)$
time (see e.g.,~\cite{Even73,Sedgewick77})

\begin{theorem}
There exists an $O(k! \cdot \text{\rm poly}(n))$-time algorithm, that
given an \scp instance with $k$ neighborly vertices, finds a
$(1+\eps)$-approximate placement.
\end{theorem}
%\begin{proof}
%The running time of the algorithm follows from the discussion above,
%the correctness of the algorithm follows directly from the fact that
%there is a correct orientation among the $k!$ possible once that
%corresponds to the optimal placement
%
%\qed
%\end{proof}

>>>>

\begin{document}
\maketitle

@@

We show that even the feasibility problem is NP-hard in general
graphs.  Hence we focus on DAGs.  We show that the problem is still
NP-hard in DAGs even for a very simple network, and even if the VNF
specification consists of only one option (i.e., the virtual DAG is a
path).
%
On the other hand, we present an FPTAS for the case where the network
is a DAG.  In addition, based on our FPTAS, we provide a randomized
algorithm for instances in which the service chain passes through a
bounded number of vertices whose degree is larger than two.
\end{abstract}

@@

We show that even feasibility is NP-hard in general networks using a
reduction from \textsc{Hamiltonian Path}.
%Hence we focus on DAGs.
We show that the problem is still NP-hard if the network is a DAG even
for a very simple network, and even if the specification consists of a
single option.  Both reductions are from \textsc{Partition}.
%
On the positive side, we present an FPTAS for the case where the
network is a DAG which is based on a non-trivial dynamic programming
algorithm.
%
Based on our FPTAS, we provide a randomized algorithm for general
networks in which there is an optimal placement whose physical path
consists of at most $k$ vertices whose degree is larger than $2$.  For
example, this can be assumed if all simple paths from $s$ to $t$
contain at most $k$ such vertices.  The algorithm computes a
$(1+\eps)$-approximate placement with high probability using $k! \log
n$ invocations of the FPTAS.
%
We also present a (deterministic) parameterized algorithm that
computes a $(1+\eps)$-approximate placement in time $O(k! \cdot
\text{poly}(n))$, where $k$ is the number of vertices in the network
whose degree is larger than $2$.

@@

\item[Virtual specification:]
%
  The description of a request for a service chain consists of a
  physical source $s \in V$, a physical destination $t \in V$, and a
  directed acyclic graph (DAG) $\calG = (\calV,\calE)$.
%
  Without loss of generality, we assume that $p(s) = p(t) = 0$.  
%
  The DAG $\calG$ has a source node $\sigma \in \calV$ and a
  destination $\tau \in \calV$. Each node $\alpha \in \calV$ represent
  a VNF that has a processing demand $p(\alpha)$.
%
  Without loss of generality, we assume that $p(\sigma) = p(\tau) =
  0$.  Each edge $\eps \in \calE$ has a bandwidth demand $b(\eps)$.

@@

In this section we consider \scp when the physical network is an
undirected graph.  Recall that even the feasibility version of \scp is
NP-hard in general networks, and therefore we focus on a special case.
%
A vertex $v \in V$ is called \emph{neighborly} if it has more than two
neighbors.
%In this section
We assume that there exists an optimal placement whose physical path
consists of at most $k$ neighborly vertices.  For example, this can be
assumed if all simple paths from $s$ to $t$ contain at most $k$
neighborly vertices.
%
We present a randomized algorithm that consists of $k!$ iterations of
the following two phases: an orientation phase that applies a random
orientation to the physical network, and an execution of the FPTAS for
DAGs given in Theorem~\ref{thm:fptas}.  The algorithm finds a
$(1-\eps)$-approximate placement with probability $(1-1/e)$, and
its running time of the algorithm is $O(t(n)k!)$, where $t(n)$ is the
time it takes to compute a placement when the physical network is a
DAG.
%
As usual, one may amplify the probability of success using repetition.

@@

\begin{figure}[t]
\centering
\scalebox{0.8}{
  \input{fig-orientation}
}
\caption{Orientation phase: on the left is the physical network, where
  only the heavy nodes are drawn.  The numbers represent the
  permutation and the dashed path represents a path of an optimal
  placement.  The orientated physical network is on the right.  It is
  enough that the internal ordering of the nodes on the dashed path is
  "correct" to ensure that the survival of the optimal placement. }
\label{fig:orientation}
\end{figure}

@@

\begin{theorem}
With probability at least $1 - e^{-1}$ the algorithm finds a
$(1+\eps)$-approximate placement after $k!$ iterations.
\end{theorem}
\begin{proof}
Consider an optimal placement with at most $k$ neighborly vertices,
and let $\pi'$ be the permutation it induces on the neighborly
vertices in the physical path.
%
If the random permutation $\pi$ orders the neighborly vertices
correctly, i.e., if $\pi(v) = \pi'(v)$, for every $v \in N$, then the
optimal placement is feasible with respect to $G_\pi$.  Hence, the
FPTAS will find a $(1+\eps)$-approximate placement in $G_\pi$.  This
placement is feasible, and therefore $(1+\eps)$-approximate placement,
with respect to $G$.
%
$\pi$ agrees with $\pi'$ with probability $1/k!$, and the probability
that $\pi$ does not agree with $\pi'$ in $k!$ iterations is $1 - (1 -
1/k!)^{k!} \leq 1 - e^{-1}$.
%
\qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{FPT Algorithm for General Networks}
We now propose an FPT algorithm for general networks that is similar to the randomized 
algorithm from Section~\ref{sec:general}.
Our FPT algorithm runs in $O(k! \cdot (A + O))$ where $k$ is the number of neighborly
vertices in the graph, $A$ is the time required to find a solution on a DAG,
and $O$ is the time required to orient the graph given a permutation on the neighborly
vertices, as discussed above.
The algorithm executes the FPTAS for DAGs given in Theorem~\ref{thm:fptas}
once for any orientation results from any of the $k!$ possible permutations and
returns the best solution.
The following theorem follows immediately.

\begin{theorem}
There is an FPT algorithm that finds a $(1+\eps)$-approximate placement after $k!$ iterations 
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Fault Tolerance}
In this section we consider the problem of a service chain placement when there 
is also demand for fault tolerance.
Formally, given a physical network, a virtual specification and a parameter 
$\phi \in \mathbb{N}$, a $\phi$-robust placement is a collection of $\phi$ placements,
$f_1,\ldots,f_\phi$,
that correspond to $\phi$ internal vertex disjoint physical paths. 
We are looking for a $\phi$-robust placement that minimize the total cost:
$$
\sum_{i = 1}^\phi \sum_j c(\alpha_j,f_i(\alpha_j))
$$
In this section we discuss how to efficiently find an optimal 
(assuming integral polynomial bounded costs) 2-robust placement.
The algorithm presented here, however, can be generalized to find an optimal $\phi$-robust
placement for any $\phi$

\subsection{Failure of the Greedy Algorithm}
Here we give a simple example why the greedy algorithm fails badly.
By greedy we mean applying the previous algorithm once to find an optimal placement, 
then removing all the vertices from the resulted path and applying the same algorithm again
to find the second placement.
Consider the physical graph as described in Figure~\ref{fig:greedy-bad} and a virtual path of
two virtual functions $\alpha$ and $\beta$.
Assume latencies, bandwidth, and processing constraints are all negligible 
and assume the following cost function 
$c(1, \alpha) = c(2, \alpha) = 1-\varepsilon, c(1, \beta) = c(2, \alpha) = 1$.
The greedy algorithm is unable to find a 2-robust placement in this case - after the 
first iteration there is no path from $s$ to $t$ - while, clearly there is a 2-robust
placement that costs $4 - 2 \epsilon$.   
\begin{figure}
\caption{
\label{fig:greedy-bad}
}
\begin{center}
\input{fig-alg-bad-example}
\end{center}
\end{figure}

\subsection{Fault Tolerance on Layered Graph}
We first assume that the input graph is a directed layered graph, i.e.
$G = (\bigcup_{i = 1}^l L_i, A)$ where 
$A \subseteq \bigcup_{i = 1}^{l - 1} L_i \times L_{i+1}$.
We assume that the number of vertices in each layer is at least two, or otherwise a 2-robust 
placement does not exist.
The algorithm uses dynamic programming.
We find the optimal 2-robust placement for every  
combination of two physical vertices from the same layer, 
and every combination of two virtual functions.
Formally, given the input $(G = (V, E), \calG = (\calV, \calE), p, b, s, t, c)$
We define $ft(v_1, v_2, \alpha_1, \alpha_2)$ to be the optimal 2-robust placement of 2 chains
that start at $\sigma$ and end at $\alpha_1, \alpha_2$ respectively into 
(internal vertex disjoint) physical paths that start at $s$ and end at $v_1,v_2$ respectively.
One can verified that the following recursive formula holds:
$$
ft(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) = \min
\begin{cases}
nn(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2)
\\
ny(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2)
\\
yn(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2)
\\
yy(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) 
\end{cases}
$$

where
$$
\begin{array}{l}
nn(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) = 
\displaystyle\smashoperator{\min_{\substack{\\ u_1, u_2: \\ b(u_1, v_1) \geq b(\gamma_1, \beta_1) \\ b(u_2, v_2) \geq b(\gamma_2, \beta_2)}}}
ft(u_1, u_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2)
\\
ny(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) = 
\displaystyle\smashoperator{\min_{
	\substack{
	u_1, u_2: 
	\\ 
	b(u_1, v_1) \geq b(\gamma_1, \beta_1) 
	\\ 
	b(u_2, v_2) \geq b(\gamma_2, \beta_2)
	\\
	\gamma'_2\beta'_2 \in \calE:
	\\
	\beta'_2 \prec \gamma_2 
	}}}
ft(u_1, u_2, \alpha_1, \gamma_2, \gamma_1\beta_1, \gamma'_2\beta'_2)
\\
yn(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) = 
\displaystyle\smashoperator{\min_{
	\substack{
	u_1, u_2: 
	\\ 
	b(u_1, v_1) \geq b(\gamma_1, \beta_1) 
	\\ 
	b(u_2, v_2) \geq b(\gamma_2, \beta_2)
	\\
	\gamma'_1\beta'_1 \in \calE:
	\\
	\beta'_1 \prec \gamma_1 
	}}}
ft(u_1, u_2, \gamma_1, \alpha_2, \gamma'_1\beta'_1, \gamma_2\beta_2)
\\
yy(v_1, v_2, \alpha_1, \alpha_2, \gamma_1\beta_1, \gamma_2\beta_2) = 
\displaystyle\smashoperator{\min_{
	\substack{
	u_1, u_2: 
	\\ 
	b(u_1, v_1) \geq b(\gamma_1, \beta_1) 
	\\ 
	b(u_2, v_2) \geq b(\gamma_2, \beta_2)
	\\
	\gamma'_1\beta'_1, \gamma'_2\beta'_2 \in \calE:
	\\
	\beta'_1 \prec \gamma_1, \beta'_2 \prec \gamma_2 
	}}}
ft(u_1, u_2, \gamma_1, \gamma_2, \gamma'_1\beta'_1, \gamma'_2\beta'_2)
\end{array}
$$

\begin{theorem}
There is an algorithm that, on a layered graph, finds an optimal 2-robust placement in 
polynomial time. 
\end{theorem}

\begin{proof}
One can verify, that using dynamic programming, the above recursive can be solved in polynomial time,
using backtracking we can also find the optimal solution.
\end{proof}

We note that the above algorithm does not take into account the latency constraint.
It is possible, however, given a latency constraint, $L$, to modify the algorithm to return
an optimal solution such that the total latency of the two placements is no more than $2L$.
Such a solution is guaranteed to be optimal in terms of costs.
It is also easy to see that in such solution at least one of the placements has lower latency 
than $L$ and the other one has latency that is at most $2L$.

\subsection{Layered Graph Transformation}
Given a directed acyclic graph $G = (V, A)$, 
and let $\pi:V \to [n]$ be an arbitrary topological sorting of $G$ 
we transform it into a layered graph $G'$ by replacing every edge $(u, v)$ with a directed path of 
length (number of arcs) $\pi(v) - \pi(u)$ from $u$ to $v$.
We set the processing capacity of each internal vertex on this path to be zero,
and the bandwidth of the arcs along the path to be $b(u,v)$.
It is easy to see that the resulting graph is indeed a layered graph and that every placement
on $G$ corresponds to a placement on $G'$ with the same cost.
Figure~\ref{fig:layer-transform} depicts this process.
\begin{figure}
\captionsetup{singlelinecheck=off}
\caption[layered]{foo
\label{fig:layer-transform}
\begin{enumerate}
\item A directed acyclic graph that is not a layered graph
  \item A topological sorting of the above graph
  \item A layered graph obtained by replacing crossing arcs with simple paths 
\end{enumerate}}
\begin{center}
\input{layer-transform}
\end{center}
\end{figure}
